A new method to estimate dark matter halo concentrations

### 2015-01-21 paper/article.tex
intro de sec. resultados. subsubsec impact of particle number

We apply our method to two different halo samples. The first one is a
mock sample to test the algorithm and asses the impact of the number
of computational particles on the concentration values. The second
sample comes from a publicly available N-body cosmological
simulation. These results for these sample can be then compared
against other methods to find the concentration and feed the
discussion about the implications of our method.

\subsection{Mock Halos}

First, we test our method using mock halos with known values for the
concentration. The method we use to generate the halos is based on the integrated
mass profile. Given the number of particles $n$ and the concentration
$c$ we define the mass element as $\delta m = 1/n$ (this corresponds
to the mass of each particle such that the total mass is one). Then
for each number $k$ from $1$ to $n$ we find the value of $r$ such that
the difference
%
\begin{equation}
m(<r;c)-k \cdot \delta m
\end{equation}
%
is zero using Ridders' method. This value of $r$ is the radius of
the $k$th particle of the generated halo, then polar and azimuthal
angles $\theta$ and $\phi$ are randomly generated. Finally these three
coordinates are transformed into cartesian coordinates
$(r,\theta,\phi) \rightarrow (x,y,z)$. This process is repeated $n$
times in order to generate the $x,y,z$ coordinates for each particle.

We generate $100$ mock halos with concentration values randomly
placed in the range $1<c<20$. Each one of the halos is generated with
four different total particle numbers: $20$, $200$, $2000$ and
$20000$. This gives us a total of $500$ halos.

\begin{figure}
\begin{center}
  \includegraphics[width=0.48\textwidth]{percentual_diff.pdf}
\end{center}
\caption{Cumulative distribution of the fractional difference between
  the input concentration in the mock halo generator, $c_{in}$ and the
  measurement by our MCMC code, $c_{out}$. Each curve corresponds to
  halos generated with a different number of particles, $N$.
    \label{fig:results_mocks}}
\end{figure}

In Figure \ref{fig:results_mocks} we show the integrated distribution
percentual difference between the concentration measured with our
fitting method $c_{out}$ in comparison with the concentration used to
generate the mock halos, $c_{in}$, $D=(c_{in}-c_{out})/c_{in}$. Then we can see that as the number of particles increases, the curve becomes more pronounced at 0. Showing that for most of the halos $c_{in}$ is very similar to $c_{out}$.

>>>>

In this Section we present the results of applying our method on two
different halo samples.

The first sample is composed by mock halos generated to have known
concentration values in perfect spherical symmetry following an NFW
profile. 
We use this sample to check that we can recover the expected values
but also gauge the impact of the number of particles on the outcomes
and the difference with respect to the two other fitting methods.

The second sample comes from a publicly available N-body cosmological
simulation. 
From this sample we quantify again the differences between all the
methods we have to fit the data.
We also estimate the possible impact of the different methods in
estimating the mass-concentration relationship from simulations.

\subsection{Tests on Mock Halos}

The method we use to generate the halos is based on the integrated
mass profile. 
We start by fixing the desired concentration $c$ and total number of
particles $N$ in the mock halo 
With this values we define the mass element as $\delta m = 1/M$, corresponding
to the mass of each particle such that the total mass is one. 
Then for each particle $i=1,\ldots,N$, we find the value of $r_i$ such that
the difference
%
\begin{equation}
m(<r_i;c) - i \cdot \delta m
\end{equation}
%
is zero using Ridders' method.

The value of $r_i$ is the radius of the $i$-th particle of the mock
halo.
Then we generate random polar and azimuthal angles $\theta$ and $\phi$
for each particle to ensure spherical symmetry. 
Finally these three spherical coordinates are transformed into cartesian coordinates
$(r,\theta,\phi) \rightarrow (x,y,z)$.

We generate in total $400$ mock halos splitted into four different
groups of $100$ halos each. 
The four groups differ in the total number of particles for their halos:
$20$, $200$, $2000$ and $20000$.
Inside each group the halos have random concentration values in
the range $1<c<20$ with a flat distribution. 
For all these halos with find the concentration values using the
density, velocity and mass methods described in the previous
section. We quantify the difference between the expected $c_{in}$
and obtained $c_{out}$ values by $D=(c_{in}-c_{out})/c_{in}$.

\subsubsection{The impact of particle number}

\begin{figure}
\begin{center}
  \includegraphics[width=0.50\textwidth]{percentual_diff.pdf}
\end{center}
\caption{Cumulative distribution of the fractional difference, $D$, between
  the input concentration in the mock halo generator, $c_{in}$ and the
  measurement by our MCMC code, $c_{out}$. Each curve corresponds to
  halos generated with a different number of particles, $N$.
    \label{fig:results_mocks}}
\end{figure}

Figure \ref{fig:results_mocks} shows the integrated distribution for
$D$ for the fits using our method, splitted into four different groups
according the particle number. 
From this Figure the first immediate
conclusion is that increasing the number of particles increases the
chances to recover the input values.

We believe that the main effect that contributes to this trend
is that the particle that our algorithm finds to be the halo center
(where the potential is minimum) gets closer to the original
geometrical center (where no particle sits by construction) used to
generate the halo.  Poissonian noise makes this center fluctuate,
changing the numerical radial profile from the analytical input.

For particle number of $20$ the offset betwen the input and output
concentration can be as large as $20\%$, with a slight bias around
$-0.05\%$, i.e. the output concentration is biased towards lower
values than the input.
For particle numbers of $2000$ most of the offsets fall below $5\%$,
with a clear peak around $0\%$ indicating that any appreciable bias is
absent.

\subsubsection{The impact of the input concentration}

Hace falta hacer una figura para este caso.
